A data-driven, pixel-level model to improve the precision of \Kepler\ photometry

### 2013-10-15 documents/plm.tex
crazy generalities and philosophy

The point of these models is to be \emph{flexible},
  so the usual approach is to make the input data set very large,
  and the number of parameters large,
  often as large as---or larger than---the target data available.
Such fits require \emph{regularization} to break degeneracies
  and control ill-constrained parameters.
These regularizations do for optimizations what priors do for posterior probability inferences;
  they express the desired behavior of the fit in the absence of data
  or along directions in parameter space in which the data are not constraining.
General regularizations or priors will break the convexity of linear models;
  if convexity is to be maintained, these regularizations also have to be quadratic,
  or else one of a small number of other known forms (one of which is L1-regularization).
Given these considerations, it makes sense to try to build our pixel-level model
  using a linear model with a quadratic objective function and a quadratic regularization.

@@

The third general idea to mention is that we need to control for over-fitting.
That is, once you have a flexible-enough model you can in principle fit \emph{anything},
  whether it was caused by the spacecraft, intrinsic stellar variability, or a transiting exoplanet.
How do you prevent a very flexible model from using small noise-induced fluctuations in all the input data
  from being carefully combined linearly into models for every nook and cranny in the target data?
In most projects in astrophysics, over-fitting is controlled for by limiting model freedom.
The model is restricted in the number of parameters
  (as in ``you can't have more parameters than data points'')
  or by limiting the dimensionality
  (as in principal components analysis)
  or by applying strong priors
  (as with smoothness priors, that effectively reduce freedom without explicitly reducing the number of parameters).
In each case, the restriction on the model can be set with tools like cross-validation,
  the fully marginalized likelihood (Bayes factor or evidence),
  chi-squared statistical criteria,
  or intuition or heuristics.
Here we take a different and more general approach, which is to use a train-and-test framework.

@@

Fourth idea: predict; residuals are our signal...

Fifth idea: no way to know if we are \emph{right}...

\section{Specifics: PLM and model hyper-parameters}

>>>>

The point of these models is to be \emph{flexible},
  so the usual approach is to make the input data set very large,
  and the number of parameters large,
  often as large as---or larger than---the target data available.
Such fits require \emph{regularization} to break degeneracies
  and control ill-constrained parameters.
These regularizations do for optimizations what priors do for posterior probability inferences;
  they express the desired behavior of the fit in the absence of data
  or along directions in parameter space in which the data are not constraining.
General regularizations or priors will break the convexity of linear models;
  if convexity is to be maintained, these regularizations also have to be quadratic,
  or else one of a small number of other known forms (one of which is L1-regularization).
Given these considerations, it makes sense to try to build our pixel-level model
  using a linear model with a quadratic objective function and a quadratic regularization;
  this is what \name\ will be.

@@

The third general idea is that we need to control for over-fitting.
That is, once you have a flexible-enough model you can in principle fit \emph{anything},
  whether it was caused by the spacecraft, intrinsic stellar variability, or a transiting exoplanet.
How do you prevent a very flexible model from using small noise-induced fluctuations in all the input data
  from being carefully combined linearly into models for every nook and cranny in the target data?
In most projects in astrophysics, over-fitting is controlled for by limiting model freedom.
The model is restricted in the number of parameters
  (as in ``you can't have more parameters than data points'')
  or by limiting the dimensionality
  (as in principal components analysis)
  or by applying strong priors
  (as with smoothness priors, that effectively reduce freedom without explicitly reducing the number of parameters).
In each case, the restriction on the model can be set with tools like cross-validation,
  the fully marginalized likelihood (Bayes factor or evidence),
  chi-squared statistical criteria,
  or intuition or heuristics.
Here we take a different and more general approach, which is to use a train-and-test framework.

@@

The fourth general idea involved in this kind of data-driven modeling is that the models aren't \emph{interpretable}.
This means, in particular, that although the model might do a good job of \emph{predicting} the target pixels,
  using a linear (or more complex) combination ofthe input pixel data,
  it won't deliver anything that can be unambiguously interpreted as the \emph{flux} of the star in question,
  or any other signal we care about.
The data-driven model \emph{effectively} describes the pointing, point-spread function, and flat-field
  of the telescope and camera,
  plus the variability of every star and every exoplanet transit,
  but it does so without ever \emph{explicitly} creating any of those objects.
We have to make some kind of heuristic or interpretive move to extract from the data-driven model the quanitites of interest.

In the context of this project, the quantities of interest are the light-curves
  (photometry as a function of time or apparent flux as a function of time)
  for every star.
What we get, for every star, from the data-driven model, is a train-and-test prediction
  for the brightness of every pixel in the patch of the detector relevant to the star,
  along with the residual, or the difference between the measured read-out pixel value and the prediction (data minus model).
What will be called the ``\name\ photometry'' in what follows is constructed from what will be called the ``\name\ images''.
The \name\ photometry is constructed from the \name\ images just as the \Kepler\ official SAP photometry is constructed from the XXX images;
  that is, by summing pre-defined sets of pixels (with unit weights; see above and below for criticism of this).
The \name\ images are constructed from the data-driven model predictions by taking,
  for each pixel in each contiguous data segment (see below),
  the \emph{mean} predicted pixel value (mean over time),
  and adding to it the residuals (data minus model) away from the data-driven model prediction.
These \name\ images are ``calibrated'' images,
  in that the data-driven model has been used to fit out all the variability it can,
  subject to the restrictions imposed by its finite freedom and the train-and-test specification.
The signal---in which exoplanet transits will be found---is fundamentally the observed residual away from this prediction,
  but it has to be added back into the mean prediction to have the right ``baseline'' or zero level.
In short, \name\ will produce prediction images, residual images, \name\ images (mean prediction plus residual),
  and aperture-based \name\ photometry analogous to the SAP photometry.

Finally, the fifth general idea is that there is no objective \emph{ground truth} against which we can tell
  that any particular data-driven model is better or worse than any other.
This problem is a problem for \name, and for the \Kepler\ PDC, and any other data-driven calibration models.
One might think that the ``best'' model is the one that predicts data with lowest variance;
  this would be true if all models adhered to the same train-and-test framework, which they don't.
Besides, a model that can predict not just the spacecraft-induced variability,
  but also variability caused by exoplanet transits of interest,
  will effectively over-fit and distort the most important information in the light-curves.
That is, what is considered best for modeling the data depends on the objectives of the user.
For us, who are interested (in the long term) in finding and characterizing Earth analogs,
  the ``best'' data-driven model is the one that produces the most success in finding and characterizing Earth analogs!
What we will show in what follows is that the \name\ photometry does not distort
  artificially exoplanet transit signals injected into real \Kepler\ pixel-level data.
We can also show that it produces stellar light-curves that look sensible for known variable stars.
We cannot yet show that it is optimal for our long-term goals,
  nor can we show that it produces images or light-curves that are more accurately calibrated;
  we will only show promising properties of the \name\ outputs.
In the end, the value of the \name\ will be demonstrated by the scientific projects it enables.

\section{Specifics: \name\ specification and hyper-parameters}
